\texttt{davos}: a Python package ``smuggler'' for constructing lightweight reproducible notebooks

### 2023-04-17 paper/main.tex
added pitfalls and limitations section

@@

>>>>

Although existing notebooks \textit{can} incorporate system calls that install
project requirements, handling project requirements in the general case is
non-trivial (e.g., see Fig.~\ref{fig:flow-chart}). Further, \texttt{davos}
incorporates its own virtual environment system that isolates
notebook-installed packages from the runtime environment
(Sec.~\ref{subsec:projects}). In many setups this feature can eliminate the
need to set up a seperate virtual environment or container (e.g., in
conjunction with a \texttt{requirements.txt}, \texttt{project.toml}, or
\texttt{environment.yml} file specifying the project's dependencies).

@@

\subsection{Pitfalls and limitations}

While \texttt{davos} enables developers to conveniently specify all project
dependencies, there are some edge cases and limitations that are worth
considering. Many Python packages include (in their \texttt{setup} options)
additional dependencies that often carry their own version specifications.
Although \texttt{davos} will check that the correct version of the requested
top-level package is installed and imported into the workspace, the version
numbers of any \textit{dependencies} of the request package are \textit{not}
checked. In principle, this could lead to unexpected behavior, for example if a
given package's dependencies (or dependencies of those dependencies, etc.) were
left under-specified. A developer could mitigate this by explicitly smuggling
exact version numbers of \textit{every} project dependency (e.g., obtained via
\texttt{pip freeze}). However, for projects where the versions of dependencies
of \texttt{smuggle}d packages also need to be precisely controlled, a lockfile
or a \texttt{requirements.txt} file produced by \texttt{pip freeze} (i.e.,
explicitly specifying \textit{all} package's version numbers) may provide a
more comprehensive alternative to \texttt{davos}.

Reproducibility is not solely about dependency management. In addition to
ensuring that project dependencies are satisfied, the user running a given
notebook must also execute the code in the indicated order. For example, the
cells in a notebook may be manually run out of order. If different cells in a
\texttt{davos}-enhanced notebook made use of different versions of the same
package, this could result in \textit{more} confusion or \textit{greater}
replication failure rates relative to standard Jupyter notebooks. Therefore an
important consideration when using \texttt{davos} is that it is perhaps even more
important to execute notebook cells in order than would be the case in the
standard (non-\texttt{davos}) setup.  We suggest that developers who wish to use
\texttt{davos} include notes regarding which cells must be executed in sequence.

As of this writing, \texttt{davos} can install packages using \textt{pip}, but
not other standard Python package management systems such as \texttt{conda}.
Therefore packages that are not installable via \texttt{pip} are currently
unsupported by \texttt{davos}.  We anticipate adding support for other
package management systems, including \texttt{conda} in a future release.
